---
title: "Titanite Petrochronology and Geochemistry"
subtitle: "Porphyry Copper Deposits in Sonora, Mexico"
author: "Dylan Enwia"
institute: "University of Illinois at Chicago"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(broom)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  dpi = 300,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "80%",
  message = FALSE,
  warning = FALSE
)
```

```{r load-data, include=FALSE}
Titanite <- read.csv("Titanite.csv")

Titanite$Sample_ID <- factor(
  Titanite$Sample_ID,
  levels = c("25CU01", "25SV1-1", "25CV02", "25LC2", "25LV1")
)

Titanite <- Titanite %>%
  filter(!is.na(Sample_ID))

model1 <- lm(Temperature_C ~ Age_Ma, data = Titanite)
model2 <- lm(Eu_Anomaly ~ Age_Ma, data = Titanite)
model3 <- lm(Ce_Anomaly ~ Age_Ma, data = Titanite)
```

```{r style, include=FALSE}
style_xaringan(
  title_slide_background_color = "#1f2933",
  title_slide_text_color = "white",
  header_color = "#0f172a",
  text_color = "#1f2933",
  link_color = "#2563eb"
)
```


class: center, middle

## Research Question

How does titanite age relate to crystallization temperature and REE anomalies?

---

# Why Titanite?

Titanite is useful because it can record both age and geochemical information from igneous and hydrothermal systems.

.pull-left[
## Petrochronology

- U-Pb dating gives titanite crystallization age
- Age helps separate different magmatic events
- Useful for comparing older vs. younger titanite populations
]

.pull-right[
## Geochemistry

- Zr-in-titanite gives temperature estimates
- Eu anomaly tracks magma evolution
- Ce anomaly reflects oxidation state
]

---

# Data & Variables

The dataset includes 279 titanite analyses from five Sonora samples:

.pull-left[
## Samples

- Cuitaca (25CU01)
- Cerro Verde (25CV02)
- Cerro Saupchi (25LC2)
- Puerto La Valdeza (25LV1)
- Suaqui Verde (25SV1-1)
]

.pull-right[
## Variables

**Predictor variable**

- `Age_Ma`

**Response variables**

- `Temperature_C`
- `Eu_Anomaly`
- `Ce_Anomaly`
]

---

# Summary of Dataset

```{r summary-table, echo=FALSE}
Titanite %>%
  group_by(Sample_ID) %>%
  summarize(
    Count = n(),
    Mean_Age = mean(Age_Ma, na.rm = TRUE),
    Mean_Temp = mean(Temperature_C, na.rm = TRUE),
    Mean_Eu = mean(Eu_Anomaly, na.rm = TRUE),
    Mean_Ce = mean(Ce_Anomaly, na.rm = TRUE)
  ) %>%
  kable(format = "html", digits = 2)
```

---

# Age vs. Temperature Plot

.center[
```{r age-temperature, echo=FALSE, fig.width=6, out.width="75%"}
ggplot(Titanite, aes(x = Age_Ma, y = Temperature_C, color = Sample_ID)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  theme_minimal() + 
  labs(
    title = "Titanite Age vs. Zr-in-Titanite Temperature",
    x = "Titanite U-Pb Age (Ma)",
    y = "Zr-in-Titanite Temperature (°C)",
    color = "Sample ID"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )
```
]

---

# Age vs. Eu Anomaly Plot

.center[
```{r age-eu-anomaly, echo=FALSE, fig.width=6, out.width="75%"}
ggplot(Titanite, aes(x = Age_Ma, y = Eu_Anomaly, color = Sample_ID)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  theme_minimal() +
  labs(
    title = "Titanite Age vs. Eu/Eu*",
    x = "Titanite U-Pb Age (Ma)",
    y = "Eu/Eu*",
    color = "Sample ID"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )
```
]

---

# Age vs. Ce Anomaly

.center[
```{r age-ce-anomaly, echo=FALSE, fig.width=6, out.width="75%"}
ggplot(Titanite, aes(x = Age_Ma, y = Ce_Anomaly, color = Sample_ID)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  theme_minimal() +
  labs(
    title = "Titanite Age vs. Ce Anomaly",
    x = "Titanite U-Pb Age (Ma)",
    y = "Ce Anomaly",
    color = "Sample ID"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom"
  )
```
]

---

# Linear Regression Results

```{r regression-results, echo=FALSE}
model_ci <- bind_rows(
  Temperature = tidy(model1, conf.int = TRUE, conf.level = 0.95),
  Eu_Anomaly = tidy(model2, conf.int = TRUE, conf.level = 0.95),
  Ce_Anomaly = tidy(model3, conf.int = TRUE, conf.level = 0.95),
  .id = "Response"
) %>%
  filter(term == "Age_Ma") %>%
  select(Response, estimate, conf.low, conf.high, p.value)

model_fit <- bind_rows(
  Temperature = glance(model1),
  Eu_Anomaly = glance(model2),
  Ce_Anomaly = glance(model3),
  .id = "Response"
) %>%
  select(Response, r.squared, adj.r.squared)

model_ci %>%
  left_join(model_fit, by = "Response") %>%
  select(Response, estimate, conf.low, conf.high, r.squared, adj.r.squared, p.value) %>%
  kable(
    format = "html",
    digits = 4,
    col.names = c(
      "Response",
      "Slope Estimate",
      "95% CI Lower",
      "95% CI Upper",
      "R²",
      "Adjusted R²",
      "p-value"
    )
  )
```

- These models test whether titanite age predicts each response variable.

---


# Conclusions

- Titanite provides both age and geochemical information.
- Age can be tested as a predictor of temperature, Eu/Eu*, and Ce anomaly.
- Eu/Eu* may reflect magma evolution, plagioclase effects, or redox conditions.
- Ce anomaly may relate to oxidation state, which is important for porphyry copper fertility.
- These results help connect titanite geochemistry to the larger question of magmatic fertility in Sonora.

---

class: center, middle

#Thank you

<span style="font-size: 60px; font-weight: bold;">Questions?</span>